---
output: rmarkdown::github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(collapse=TRUE, comment="##", fig.retina=2, fig.path = "README_figs/README-")
```

Roadmap:

predict electricity load* with and without wind power using linear and non-linear methods,

report results/conclusions

-data exploration,data visualizations (make heatmap)

*predict 24 hour load using past history of demands and identify how wind energy affects energy forecasting
# Electricity Load Prediction in Texas



## Introduction

How do electric companies know how much power they have to generate? 


But why is it important to predict hourly demand for electricity at least a day in advance? You need to know much generators needs to be on to meet the expected demand and turning on a generator requires time.



## Libraries/packages we will be using
```{r libraries, message=FALSE, warning=FALSE}
new_cran_packages <- c("ggplot2", "caret","stringr")
existing_packages <- installed.packages()[,"Package"]
missing_packages <- new_cran_packages[!(new_cran_packages %in% existing_packages)]
if(length(missing_packages)){
    install.packages(missing_packages)
}

library(ggplot2)
library(stringr)
library(caret)
```


## Load the ERCOT 2018 data
```{r dataset, echo=FALSE}
dfDemand = read.csv("ERCOT_2018_Hourly_Wind_Output.csv")
demands = dfDemand$ERCOT.Load..MW
numberOfDays = length(demands)/24
```


Let's see how does load vary over the year in Texas.
```{r electricity graph, echo=FALSE}
days = vector(length = numberOfDays * 24)
for (hour in seq_len(numberOfDays * 24)) {
  days[hour] = hour / 24 
}

ggplot(dfDemand, aes(x = days, y = demands)) + geom_line(color = "dodgerblue") + 
  labs(title = "Texas Electricity Demands in 2018", x = "Days in 2018", y = "Net Demand of Texas (in MW)") +
  theme_minimal()
```


## Wind data

```{r wind output graph, echo=FALSE}
dfDemand$Windless_Load = dfDemand$ERCOT.Load..MW - dfDemand$Total.Wind.Output..MW
windDemand = dfDemand$Windless_Load
windOutput = dfDemand$Total.Wind.Output..MW

ggplot(dfDemand, aes(x = days, y = windOutput)) + geom_line(color = "orange") + 
  labs(title = "Texas Wind Power Output in 2018", x = "Days in 2018", y = "MegaWatts") + theme_minimal()
```

Wind Power looks very sporadic while electricity demands seems to have a trend.

```{r January HeatMap, echo=FALSE}
monthOfJan = dfDemand[grep("Jan", dfDemand$Date), ]
monthOfJan$Day = str_split_fixed(monthOfJan$誰..time.date.stamp, " ", 2)[,1]
monthOfJan$Hour = str_split_fixed(monthOfJan$誰..time.date.stamp, " ", 2)[,2]
monthOfJan$Day = as.Date(monthOfJan$Day, "%m/%d/%Y")
monthOfJan$Hour = as.numeric(format(strptime(monthOfJan$Hour,"%H:%M:%S"),'%H'))

#monthOfJan = monthOfJan %>% 
 # mutate(weekday = weekdays(Day)) %>%
  #mutate(weekday = factor(weekday, levels = c("Sunday", "Monday", "Tuesday", "Wednesday",
   #                                           "Thursday", "Friday", "Saturday")))

theme_heatmap = theme_light() +
  theme(panel.grid = element_blank(),
        panel.border = element_blank(),
        plot.title = element_text(face = "bold", size = 11, hjust = 0.5),
        axis.ticks = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 10),
        axis.text.x = element_text(size = 10, angle = 90, hjust = 1),
        axis.text.y = element_text(size = 8))

ggJan = ggplot(monthOfJan, mapping = aes(x = Date, y = Hour, fill = ERCOT.Load..MW)) +
  geom_tile(colour = "white") + 
  scale_fill_gradient(low = "#f4fff6", high= "#0bb730") +
  scale_y_reverse(breaks=c(23:0), labels=c(23:0), expand = c(0,0)) +
  scale_x_discrete(expand = c(0,0), position = "top") +
  labs(title = "Electricity Demand Profile in January 2018", y = "Hour of Day", fill = "Load (in MW)") + 
  theme_heatmap 

print(ggJan)

```
```{r July HeatMap, echo=FALSE}
monthOfJuly <- dfDemand[grep("Jul", dfDemand$Date), ]
monthOfJuly$Day = str_split_fixed(monthOfJuly$誰..time.date.stamp, " ", 2)[,1]
monthOfJuly$Hour = str_split_fixed(monthOfJuly$誰..time.date.stamp, " ", 2)[,2]
monthOfJuly$Day = as.Date(monthOfJuly$Day, "%m/%d/%Y")
monthOfJuly$Hour = as.numeric(format(strptime(monthOfJuly$Hour,"%H:%M:%S"),'%H'))

ggJuly = ggplot(monthOfJuly, mapping = aes(x = Date, y = Hour, fill = ERCOT.Load..MW)) +
  geom_tile(colour = "white") + 
  scale_fill_gradient(low = "#f4fff6", high= "#0bb730") +
  scale_y_reverse(breaks=c(23:0), labels=c(23:0), expand = c(0,0)) +
  scale_x_discrete(expand = c(0,0), position = "top") +
  labs(title = "Electricity Demand Profile in July 2018", y = "Hour of Day", fill = "Load (in MW)") +
  theme_heatmap 

print(ggJuly)
```



## Demand Prediction Strategy and Data Aggregation

For our independent variables we will use past week, past 2 days, past 1 day to predict the electiricty demand of tomorrow. i.e days to train on -7, -2, -1


```{r organize data}
daysToTrainOn = c(-7,-2,-1)
rangeOfDays = seq(-min(daysToTrainOn), numberOfDays - 1, by = 1)
Y = NULL
for (day in rangeOfDays) {
  Y = rbind(Y, dfDemand$ERCOT.Load..MW[(day * 24): ((day + 1) * 24 - 1)])
}

X = NULL
for (day in rangeOfDays) {
  X_temp = cbind(t(dfDemand$ERCOT.Load..MW[(((day - 7)*24 +1)):((day - 7 + 1)*24)]),
            t(dfDemand$ERCOT.Load..MW[(((day - 2)*24) +1):((day - 2 + 1)*24)]),
            t(dfDemand$ERCOT.Load..MW[(((day - 1)*24) +1):((day - 1 + 1)*24)]))
  X = rbind(X, X_temp)
}
dim(X)
dim(Y)

Y = data.frame(Y)
X = data.frame(X)
```

## Windless
```{r}
#daysToTrainOn = c(-7,-2,-1)
#rangeOfDays = seq(-min(daysToTrainOn), numberOfDays - 1, by = 1)

#Y = NULL
#for (day in rangeOfDays) {
 # Y = rbind(Y, dfDemand$Windless_Load[(day * 24): ((day + 1) * 24 - 1)])
#}

#X = NULL
#for (day in rangeOfDays) {
 # X_temp = cbind(t(dfDemand$Windless_Load[(((day - 7)*24 +1)):((day - 7 + 1)*24)]),
  #          t(dfDemand$Windless_Load[(((day - 2)*24) +1):((day - 2 + 1)*24)]),
   #         t(dfDemand$Windless_Load[(((day - 1)*24) +1):((day - 1 + 1)*24)]))
  #X = rbind(X, X_temp)
#}
#dim(X)
#dim(Y)
```

After Organzing the data we will start making our train and test data.

```{r}
test_inds = createDataPartition(y = 1:nrow(Y), p = 0.2, list = F)

X_test = X[test_inds, ]; Y_test = Y[test_inds,]
X_train = X[-test_inds, ]; Y_train = Y[-test_inds,]


colnames(Y_train) = c("day0.00", "day0.01",  "day0.02", "day0.03", "day0.04", "day0.05", "day0.06", "day0.07"
                , "day0.08", "day0.09", "day0.10", "day0.11", "day0.12", "day0.13", "day0.14", "day0.15",
                 "day0.16", "day0.17", "day0.18", "day0.19", "day0.20", "day0.21", "day0.22", "day0.23")

colnames(Y_test) = c("day0.00", "day0.01",  "day0.02", "day0.03", "day0.04", "day0.05", "day0.06", "day0.07"
                , "day0.08", "day0.09", "day0.10", "day0.11", "day0.12", "day0.13", "day0.14", "day0.15",
                 "day0.16", "day0.17", "day0.18", "day0.19", "day0.20", "day0.21", "day0.22", "day0.23")

colnames(X_train) = c("day7.00", "day7.01",  "day7.02", "day7.03", "day7.04", "day7.05", "day7.06", "day7.07"
                , "day7.08", "day7.09", "day7.10", "day7.11", "day7.12", "day7.13", "day7.14", "day7.15",
                 "day7.16", "day7.17", "day7.18", "day7.19", "day7.20", "day7.21", "day7.22", "day7.23",
                
                "day2.00", "day2.01",  "day2.02", "day2.03", "day2.04", "day2.05", "day2.06", "day2.07"
                , "day2.08", "day2.09", "day2.10", "day2.11", "day2.12", "day2.13", "day2.14", "day2.15",
                 "day2.16", "day2.17", "day2.18", "day2.19", "day2.20", "day2.21", "day2.22", "day2.23",
                
                "day1.00", "day1.01",  "day1.02", "day1.03", "day1.04", "day1.05", "day1.06", "day1.07"
                , "day1.08", "day1.09", "day1.10", "day1.11", "day1.12", "day1.13", "day1.14", "day1.15",
                 "day1.16", "day1.17", "day1.18", "day1.19", "day1.20", "day1.21", "day1.22", "day1.23")

colnames(X_test) = c("day7.00", "day7.01",  "day7.02", "day7.03", "day7.04", "day7.05", "day7.06", "day7.07"
                , "day7.08", "day7.09", "day7.10", "day7.11", "day7.12", "day7.13", "day7.14", "day7.15",
                 "day7.16", "day7.17", "day7.18", "day7.19", "day7.20", "day7.21", "day7.22", "day7.23",
                
                "day2.00", "day2.01",  "day2.02", "day2.03", "day2.04", "day2.05", "day2.06", "day2.07"
                , "day2.08", "day2.09", "day2.10", "day2.11", "day2.12", "day2.13", "day2.14", "day2.15",
                 "day2.16", "day2.17", "day2.18", "day2.19", "day2.20", "day2.21", "day2.22", "day2.23",
                
                "day1.00", "day1.01",  "day1.02", "day1.03", "day1.04", "day1.05", "day1.06", "day1.07"
                , "day1.08", "day1.09", "day1.10", "day1.11", "day1.12", "day1.13", "day1.14", "day1.15",
                 "day1.16", "day1.17", "day1.18", "day1.19", "day1.20", "day1.21", "day1.22", "day1.23")
dim(X_train)
dim(X_test)
dim(Y_train)
dim(Y_test)
```

## Prediction

### linear regression 

Predict Load of tommorrow at hour i based on last 7 days at time i, last 2 days at time i and yesterday at time i where i is 0,1,2..23.
```{r}
#predict the load at each hour using linear regression

#retrieve training data at hour 0
data00 = data.frame(Y_train$day0.00,X_train$day7.00,X_train$day2.00,X_train$day1.00)

#name our variables
colnames(data00) <- c('day0','pday7','pday2','pday1')

#linear model to predict load at hour 0
mod00 = lm(day0 ~., data = data00)

#retrieve testing data at hour 0
newdat00 = data.frame(pday7 = X_test$day7.00, pday2=X_test$day2.00, pday1=X_test$day1.00)

#use our model to predict the expect load at hour 0
predict(mod00, newdata = newdat00)

# 72 numbers should be returned; 1 for each day

#repeat this for each hour, compare the results to Y_test and visualize the results

data01 = data.frame(Y_train$day0.01,X_train$day7.01,X_train$day2.01,X_train$day1.01)
colnames(data01) <- c('day0','pday7','pday2','pday1')
mod01 = lm(day0 ~., data = data01)

newdat01 = data.frame(pday7 = X_test$day7.01, pday2=X_test$day2.01, pday1=X_test$day1.01)
#predict(mod01, newdata = newdat01)

data02 = data.frame(Y_train$day0.02,X_train$day7.02,X_train$day2.02,X_train$day1.02)
colnames(data02) <- c('day0','pday7','pday2','pday1')
mod02 = lm(day0 ~., data = data02)

newdat02 = data.frame(pday7 = X_test$day7.02, pday2=X_test$day2.02, pday1=X_test$day1.02)
#predict(mod02, newdata = newdat02)

data03 = data.frame(Y_train$day0.03,X_train$day7.03,X_train$day2.03,X_train$day1.03)
colnames(data03) <- c('day0','pday7','pday2','pday1')
mod03 = lm(day0 ~., data = data03)

newdat03 = data.frame(pday7 = X_test$day7.03, pday2=X_test$day2.03, pday1=X_test$day1.03)
#predict(mod03, newdata = newdat03)

data04 = data.frame(Y_train$day0.04,X_train$day7.04,X_train$day2.04,X_train$day1.04)
colnames(data04) <- c('day0','pday7','pday2','pday1')
mod04 = lm(day0 ~., data = data04)

newdat04 = data.frame(pday7 = X_test$day7.04, pday2=X_test$day2.04, pday1=X_test$day1.04)
#predict(mod04, newdata = newdat04)

data05 = data.frame(Y_train$day0.05,X_train$day7.05,X_train$day2.05,X_train$day1.05)
colnames(data05) <- c('day0','pday7','pday2','pday1')
mod05 = lm(day0 ~., data = data05)

newdat05 = data.frame(pday7 = X_test$day7.05, pday2=X_test$day2.05, pday1=X_test$day1.05)
#predict(mod05, newdata = newdat05)

data06 = data.frame(Y_train$day0.06,X_train$day7.06,X_train$day2.06,X_train$day1.06)
colnames(data06) <- c('day0','pday7','pday2','pday1')
mod06 = lm(day0 ~., data = data06)

newdat06 = data.frame(pday7 = X_test$day7.06, pday2=X_test$day2.06, pday1=X_test$day1.06)
#predict(mod06, newdata = newdat06)

data07 = data.frame(Y_train$day0.07,X_train$day7.07,X_train$day2.07,X_train$day1.07)
colnames(data07) <- c('day0','pday7','pday2','pday1')
mod07 = lm(day0 ~., data = data07)

newdat07 = data.frame(pday7 = X_test$day7.07, pday2=X_test$day2.07, pday1=X_test$day1.07)
#predict(mod07, newdata = newdat07)

data08 = data.frame(Y_train$day0.08,X_train$day7.08,X_train$day2.08,X_train$day1.08)
colnames(data08) <- c('day0','pday7','pday2','pday1')
mod08 = lm(day0 ~., data = data08)

newdat08 = data.frame(pday7 = X_test$day7.08, pday2=X_test$day2.08, pday1=X_test$day1.08)
#predict(mod08, newdata = newdat08)

data09 = data.frame(Y_train$day0.09,X_train$day7.09,X_train$day2.09,X_train$day1.09)
colnames(data09) <- c('day0','pday7','pday2','pday1')
mod09 = lm(day0 ~., data = data09)

newdat09 = data.frame(pday7 = X_test$day7.09, pday2=X_test$day2.09, pday1=X_test$day1.09)
#predict(mod09, newdata = newdat09)

data10 = data.frame(Y_train$day0.10,X_train$day7.10,X_train$day2.10,X_train$day1.10)
colnames(data10) <- c('day0','pday7','pday2','pday1')
mod10 = lm(day0 ~., data = data10)

newdat10 = data.frame(pday7 = X_test$day7.10, pday2=X_test$day2.10, pday1=X_test$day1.10)
#predict(mod10, newdata = newdat10)

data11 = data.frame(Y_train$day0.11,X_train$day7.11,X_train$day2.11,X_train$day1.11)
colnames(data11) <- c('day0','pday7','pday2','pday1')
mod11 = lm(day0 ~., data = data11)

newdat11 = data.frame(pday7 = X_test$day7.11, pday2=X_test$day2.11, pday1=X_test$day1.11)
#predict(mod11, newdata = newdat11)

data12 = data.frame(Y_train$day0.12,X_train$day7.12,X_train$day2.12,X_train$day1.12)
colnames(data12) <- c('day0','pday7','pday2','pday1')
mod12 = lm(day0 ~., data = data12)

newdat12 = data.frame(pday7 = X_test$day7.12, pday2=X_test$day2.12, pday1=X_test$day1.12)
#predict(mod12, newdata = newdat12)

data13 = data.frame(Y_train$day0.13,X_train$day7.13,X_train$day2.13,X_train$day1.13)
colnames(data13) <- c('day0','pday7','pday2','pday1')
mod13 = lm(day0 ~., data = data13)

newdat13 = data.frame(pday7 = X_test$day7.13, pday2=X_test$day2.13, pday1=X_test$day1.13)
#predict(mod13, newdata = newdat13)

data14 = data.frame(Y_train$day0.14,X_train$day7.14,X_train$day2.14,X_train$day1.14)
colnames(data14) <- c('day0','pday7','pday2','pday1')
mod14 = lm(day0 ~., data = data14)

newdat14 = data.frame(pday7 = X_test$day7.14, pday2=X_test$day2.14, pday1=X_test$day1.14)
#predict(mod14, newdata = newdat14)

data15 = data.frame(Y_train$day0.15,X_train$day7.15,X_train$day2.15,X_train$day1.15)
colnames(data15) <- c('day0','pday7','pday2','pday1')
mod15 = lm(day0 ~., data = data15)

newdat15 = data.frame(pday7 = X_test$day7.15, pday2=X_test$day2.15, pday1=X_test$day1.15)
#predict(mod15, newdata = newdat15)

data16 = data.frame(Y_train$day0.16,X_train$day7.16,X_train$day2.16,X_train$day1.16)
colnames(data16) <- c('day0','pday7','pday2','pday1')
mod16 = lm(day0 ~., data = data16)

newdat16 = data.frame(pday7 = X_test$day7.16, pday2=X_test$day2.16, pday1=X_test$day1.16)
#predict(mod16, newdata = newdat16)

data17 = data.frame(Y_train$day0.17,X_train$day7.17,X_train$day2.17,X_train$day1.17)
colnames(data17) <- c('day0','pday7','pday2','pday1')
mod17 = lm(day0 ~., data = data17)

newdat17 = data.frame(pday7 = X_test$day7.17, pday2=X_test$day2.17, pday1=X_test$day1.17)
#predict(mod17, newdata = newdat17)

data18 = data.frame(Y_train$day0.18,X_train$day7.18,X_train$day2.18,X_train$day1.18)
colnames(data18) <- c('day0','pday7','pday2','pday1')
mod18 = lm(day0 ~., data = data18)

newdat18 = data.frame(pday7 = X_test$day7.18, pday2=X_test$day2.18, pday1=X_test$day1.18)
#predict(mod18, newdata = newdat18)

data19 = data.frame(Y_train$day0.19,X_train$day7.19,X_train$day2.19,X_train$day1.19)
colnames(data19) <- c('day0','pday7','pday2','pday1')
mod19 = lm(day0 ~., data = data19)

newdat19 = data.frame(pday7 = X_test$day7.19, pday2=X_test$day2.19, pday1=X_test$day1.19)
#predict(mod19, newdata = newdat19)

data20 = data.frame(Y_train$day0.20,X_train$day7.20,X_train$day2.20,X_train$day1.20)
colnames(data20) <- c('day0','pday7','pday2','pday1')
mod20 = lm(day0 ~., data = data20)

newdat20 = data.frame(pday7 = X_test$day7.20, pday2=X_test$day2.20, pday1=X_test$day1.20)
#predict(mod20, newdata = newdat20)

data21 = data.frame(Y_train$day0.21,X_train$day7.21,X_train$day2.21,X_train$day1.21)
colnames(data21) <- c('day0','pday7','pday2','pday1')
mod21 = lm(day0 ~., data = data21)

newdat21 = data.frame(pday7 = X_test$day7.21, pday2=X_test$day2.21, pday1=X_test$day1.21)
#predict(mod21, newdata = newdat21)

data22 = data.frame(Y_train$day0.22,X_train$day7.22,X_train$day2.22,X_train$day1.22)
colnames(data22) <- c('day0','pday7','pday2','pday1')
mod22 = lm(day0 ~., data = data22)

newdat22 = data.frame(pday7 = X_test$day7.22, pday2=X_test$day2.22, pday1=X_test$day1.22)
#predict(mod22, newdata = newdat22)

data23 = data.frame(Y_train$day0.23,X_train$day7.23,X_train$day2.23,X_train$day1.23)
colnames(data23) <- c('day0','pday7','pday2','pday1')
mod23 = lm(day0 ~., data = data23)

newdat23 = data.frame(pday7 = X_test$day7.23, pday2=X_test$day2.23, pday1=X_test$day1.23)
#predict(mod23, newdata = newdat23)
```





