---
output: rmarkdown::github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(collapse=TRUE, comment="##", fig.retina=2, fig.path = "README_figs/README-")
```

Roadmap:

predict electricity load* with and without wind power using linear and non-linear methods,

report results/conclusions

-data exploration,data visualizations (make heatmap)

*predict 24 hour load using past history of demands and identify how wind energy affects energy forecasting
# Electricity Load Prediction in Texas



## Introduction

How do electric companies know how much power they have to generate? 


But why is it important to predict hourly demand for electricity at least a day in advance? You need to know much generators needs to be on to meet the expected demand and turning on a generator requires time.



## Libraries/packages we will be using
```{r libraries, message=FALSE, warning=FALSE}
new_cran_packages <- c("ggplot2", "caret","stringr")
existing_packages <- installed.packages()[,"Package"]
missing_packages <- new_cran_packages[!(new_cran_packages %in% existing_packages)]
if(length(missing_packages)){
    install.packages(missing_packages)
}

library(ggplot2)
library(stringr)
library(caret)
```


## Load the ERCOT 2018 data
```{r dataset, echo=FALSE}
dfDemand = read.csv("ERCOT_2018_Hourly_Wind_Output.csv")
demands = dfDemand$ERCOT.Load..MW
numberOfDays = length(demands)/24
```


Let's see how does load vary over the year in Texas.
```{r electricity graph, echo=FALSE}
days = vector(length = numberOfDays * 24)
for (hour in seq_len(numberOfDays * 24)) {
  days[hour] = hour / 24 
}

ggplot(dfDemand, aes(x = days, y = demands)) + geom_line(color = "dodgerblue") + 
  labs(title = "Texas Electricity Demands in 2018", x = "Days in 2018", y = "Net Demand of Texas (in MW)") +
  theme_minimal()
```


## Wind data

```{r wind output graph, echo=FALSE}
dfDemand$Windless_Load = dfDemand$ERCOT.Load..MW - dfDemand$Total.Wind.Output..MW
windDemand = dfDemand$Windless_Load
windOutput = dfDemand$Total.Wind.Output..MW

ggplot(dfDemand, aes(x = days, y = windOutput)) + geom_line(color = "orange") + 
  labs(title = "Texas Wind Power Output in 2018", x = "Days in 2018", y = "MegaWatts") + theme_minimal()
```

Wind Power looks very sporadic while electricity demands seems to have a trend.

```{r January HeatMap, echo=FALSE}
monthOfJan = dfDemand[grep("Jan", dfDemand$Date), ]
monthOfJan$Day = str_split_fixed(monthOfJan$誰..time.date.stamp, " ", 2)[,1]
monthOfJan$Hour = str_split_fixed(monthOfJan$誰..time.date.stamp, " ", 2)[,2]
monthOfJan$Day = as.Date(monthOfJan$Day, "%m/%d/%Y")
monthOfJan$Hour = as.numeric(format(strptime(monthOfJan$Hour,"%H:%M:%S"),'%H'))

#monthOfJan = monthOfJan %>% 
 # mutate(weekday = weekdays(Day)) %>%
  #mutate(weekday = factor(weekday, levels = c("Sunday", "Monday", "Tuesday", "Wednesday",
   #                                           "Thursday", "Friday", "Saturday")))

theme_heatmap = theme_light() +
  theme(panel.grid = element_blank(),
        panel.border = element_blank(),
        plot.title = element_text(face = "bold", size = 11, hjust = 0.5),
        axis.ticks = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 10),
        axis.text.x = element_text(size = 10, angle = 90, hjust = 1),
        axis.text.y = element_text(size = 8))

ggJan = ggplot(monthOfJan, mapping = aes(x = Date, y = Hour, fill = ERCOT.Load..MW)) +
  geom_tile(colour = "white") + 
  scale_fill_gradient(low = "#f4fff6", high= "#0bb730") +
  scale_y_reverse(breaks=c(23:0), labels=c(23:0), expand = c(0,0)) +
  scale_x_discrete(expand = c(0,0), position = "top") +
  labs(title = "Electricity Demand Profile in January 2018", y = "Hour of Day", fill = "Load (in MW)") + 
  theme_heatmap 

print(ggJan)

```
```{r July HeatMap, echo=FALSE}
monthOfJuly <- dfDemand[grep("Jul", dfDemand$Date), ]
monthOfJuly$Day = str_split_fixed(monthOfJuly$誰..time.date.stamp, " ", 2)[,1]
monthOfJuly$Hour = str_split_fixed(monthOfJuly$誰..time.date.stamp, " ", 2)[,2]
monthOfJuly$Day = as.Date(monthOfJuly$Day, "%m/%d/%Y")
monthOfJuly$Hour = as.numeric(format(strptime(monthOfJuly$Hour,"%H:%M:%S"),'%H'))

ggJuly = ggplot(monthOfJuly, mapping = aes(x = Date, y = Hour, fill = ERCOT.Load..MW)) +
  geom_tile(colour = "white") + 
  scale_fill_gradient(low = "#f4fff6", high= "#0bb730") +
  scale_y_reverse(breaks=c(23:0), labels=c(23:0), expand = c(0,0)) +
  scale_x_discrete(expand = c(0,0), position = "top") +
  labs(title = "Electricity Demand Profile in July 2018", y = "Hour of Day", fill = "Load (in MW)") +
  theme_heatmap 

print(ggJuly)
```



## Demand Prediction Strategy and Data Aggregation

For our independent variables we will use past week, past 2 days, past 1 day to predict the electiricty demand of tomorrow. i.e days to train on -7, -2, -1


```{r organize data}
daysToTrainOn = c(-7,-2,-1)
rangeOfDays = seq(-min(daysToTrainOn), numberOfDays - 1, by = 1)
Y = NULL
for (day in rangeOfDays) {
  Y = rbind(Y, dfDemand$ERCOT.Load..MW[(day * 24): ((day + 1) * 24 - 1)])
}

X = NULL
for (day in rangeOfDays) {
  X_temp = cbind(t(dfDemand$ERCOT.Load..MW[(((day - 7)*24 +1)):((day - 7 + 1)*24)]),
            t(dfDemand$ERCOT.Load..MW[(((day - 2)*24) +1):((day - 2 + 1)*24)]),
            t(dfDemand$ERCOT.Load..MW[(((day - 1)*24) +1):((day - 1 + 1)*24)]))
  X = rbind(X, X_temp)
}
dim(X)
dim(Y)

Y = data.frame(Y)
X = data.frame(X)
```

## Windless
```{r}
#daysToTrainOn = c(-7,-2,-1)
#rangeOfDays = seq(-min(daysToTrainOn), numberOfDays - 1, by = 1)

#Y = NULL
#for (day in rangeOfDays) {
 # Y = rbind(Y, dfDemand$Windless_Load[(day * 24): ((day + 1) * 24 - 1)])
#}

#X = NULL
#for (day in rangeOfDays) {
 # X_temp = cbind(t(dfDemand$Windless_Load[(((day - 7)*24 +1)):((day - 7 + 1)*24)]),
  #          t(dfDemand$Windless_Load[(((day - 2)*24) +1):((day - 2 + 1)*24)]),
   #         t(dfDemand$Windless_Load[(((day - 1)*24) +1):((day - 1 + 1)*24)]))
  #X = rbind(X, X_temp)
#}
#dim(X)
#dim(Y)
```

After Organzing the data we will start making our train and test data.

```{r}
test_inds = createDataPartition(y = 1:nrow(Y), p = 0.2, list = F)

X_test = X[test_inds, ]; Y_test = Y[test_inds,]
X_train = X[-test_inds, ]; Y_train = Y[-test_inds,]


colnames(Y_train) = c("day0.00", "day0.01",  "day0.02", "day0.03", "day0.04", "day0.05", "day0.06", "day0.07"
                , "day0.08", "day0.09", "day0.10", "day0.11", "day0.12", "day0.13", "day0.14", "day0.15",
                 "day0.16", "day0.17", "day0.18", "day0.19", "day0.20", "day0.21", "day0.22", "day0.23")

colnames(Y_test) = c("day0.00", "day0.01",  "day0.02", "day0.03", "day0.04", "day0.05", "day0.06", "day0.07"
                , "day0.08", "day0.09", "day0.10", "day0.11", "day0.12", "day0.13", "day0.14", "day0.15",
                 "day0.16", "day0.17", "day0.18", "day0.19", "day0.20", "day0.21", "day0.22", "day0.23")

colnames(X_train) = c("day7.00", "day7.01",  "day7.02", "day7.03", "day7.04", "day7.05", "day7.06", "day7.07"
                , "day7.08", "day7.09", "day7.10", "day7.11", "day7.12", "day7.13", "day7.14", "day7.15",
                 "day7.16", "day7.17", "day7.18", "day7.19", "day7.20", "day7.21", "day7.22", "day7.23",
                
                "day2.00", "day2.01",  "day2.02", "day2.03", "day2.04", "day2.05", "day2.06", "day2.07"
                , "day2.08", "day2.09", "day2.10", "day2.11", "day2.12", "day2.13", "day2.14", "day2.15",
                 "day2.16", "day2.17", "day2.18", "day2.19", "day2.20", "day2.21", "day2.22", "day2.23",
                
                "day1.00", "day1.01",  "day1.02", "day1.03", "day1.04", "day1.05", "day1.06", "day1.07"
                , "day1.08", "day1.09", "day1.10", "day1.11", "day1.12", "day1.13", "day1.14", "day1.15",
                 "day1.16", "day1.17", "day1.18", "day1.19", "day1.20", "day1.21", "day1.22", "day1.23")

colnames(X_test) = c("day7.00", "day7.01",  "day7.02", "day7.03", "day7.04", "day7.05", "day7.06", "day7.07"
                , "day7.08", "day7.09", "day7.10", "day7.11", "day7.12", "day7.13", "day7.14", "day7.15",
                 "day7.16", "day7.17", "day7.18", "day7.19", "day7.20", "day7.21", "day7.22", "day7.23",
                
                "day2.00", "day2.01",  "day2.02", "day2.03", "day2.04", "day2.05", "day2.06", "day2.07"
                , "day2.08", "day2.09", "day2.10", "day2.11", "day2.12", "day2.13", "day2.14", "day2.15",
                 "day2.16", "day2.17", "day2.18", "day2.19", "day2.20", "day2.21", "day2.22", "day2.23",
                
                "day1.00", "day1.01",  "day1.02", "day1.03", "day1.04", "day1.05", "day1.06", "day1.07"
                , "day1.08", "day1.09", "day1.10", "day1.11", "day1.12", "day1.13", "day1.14", "day1.15",
                 "day1.16", "day1.17", "day1.18", "day1.19", "day1.20", "day1.21", "day1.22", "day1.23")
dim(X_train)
dim(X_test)
dim(Y_train)
dim(Y_test)
```

## Prediction

### linear regression 

Predict Load of tommorrow at hour i based on last 7 days at time i, last 2 days at time i and yesterday at time i where i is 0,1,2..23.
```{r}
<<<<<<< HEAD
lm(Y_train$day0.00 ~ X_train$day7.00 + X_train$day2.00 + X_train$day1.00)
lm(Y_train$day0.01 ~ X_train$day7.01 + X_train$day2.01 + X_train$day1.01)
lm(Y_train$day0.02 ~ X_train$day7.02 + X_train$day2.02 + X_train$day1.02)
lm(Y_train$day0.03 ~ X_train$day7.03 + X_train$day2.03 + X_train$day1.03)
lm(Y_train$day0.04 ~ X_train$day7.04 + X_train$day2.04 + X_train$day1.04)
lm(Y_train$day0.05 ~ X_train$day7.05 + X_train$day2.05 + X_train$day1.05)
lm(Y_train$day0.06 ~ X_train$day7.06 + X_train$day2.06 + X_train$day1.06)
lm(Y_train$day0.07 ~ X_train$day7.07 + X_train$day2.07 + X_train$day1.07)
lm(Y_train$day0.08 ~ X_train$day7.08 + X_train$day2.08 + X_train$day1.08)
lm(Y_train$day0.09 ~ X_train$day7.09 + X_train$day2.09 + X_train$day1.09)
lm(Y_train$day0.10 ~ X_train$day7.10 + X_train$day2.10 + X_train$day1.10)
lm(Y_train$day0.11 ~ X_train$day7.11 + X_train$day2.11 + X_train$day1.11)
lm(Y_train$day0.12 ~ X_train$day7.12 + X_train$day2.12 + X_train$day1.12)
lm(Y_train$day0.13 ~ X_train$day7.13 + X_train$day2.13 + X_train$day1.13)
lm(Y_train$day0.14 ~ X_train$day7.14 + X_train$day2.14 + X_train$day1.14)
lm(Y_train$day0.15 ~ X_train$day7.15 + X_train$day2.15 + X_train$day1.15)
lm(Y_train$day0.16 ~ X_train$day7.16 + X_train$day2.16 + X_train$day1.16)
lm(Y_train$day0.17 ~ X_train$day7.17 + X_train$day2.17 + X_train$day1.17)
lm(Y_train$day0.18 ~ X_train$day7.18 + X_train$day2.18 + X_train$day1.18)
lm(Y_train$day0.19 ~ X_train$day7.19 + X_train$day2.19 + X_train$day1.19)
lm(Y_train$day0.20 ~ X_train$day7.20 + X_train$day2.20 + X_train$day1.20)
lm(Y_train$day0.21 ~ X_train$day7.21 + X_train$day2.21 + X_train$day1.21)
lm(Y_train$day0.22 ~ X_train$day7.22 + X_train$day2.22 + X_train$day1.22)
lm(Y_train$day0.23 ~ X_train$day7.23 + X_train$day2.23 + X_train$day1.23)
=======
data00 = data.frame(Y_train$day0.00,X_train$day7.00,X_train$day2.00,X_train$day1.00)
colnames(data00) <- c('day0','pday7','pday2','pday1')
mod00 = lm(day0 ~., data = data00)

newdat = data.frame(pday7 = X_test$day7.00, pday2=X_test$day2.00, pday1=X_test$day1.00)
predict(mod00, newdata = newdat)

#repeat this for the other times, then visualize the results

>>>>>>> 079eafdeecd59329840a23507c5a76a08a5e820a
```





