---
output: rmarkdown::github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(collapse=TRUE, comment="##", fig.retina=2, fig.path = "README_figs/README-")
```

Roadmap:

-data exploration

# Electricity Load Prediction in Texas

predict 24 hour load using past history of demands with statistical methods

identify how wind energy affects energy forecasting

data visualizations

## Introduction

How do electric companies know how much power they have to generate? 


But why is it important to predict hourly demand for electricity at least a day in advance? You need to know much generators needs to be on to meet the expected demand and turning on a generator requires time.



## Libraries/packages we will be using
```{r}
library(ggplot2)
```


## Load the ERCOT 2018 data
```{r}
dfDemand = read.csv("ERCOT_2018_Hourly_Wind_Output.csv")
demands = dfDemand$ERCOT.Load..MW
numberOfDays = length(demands)/24
```


Let's see how does load vary over the year in Texas.
```{r}
days = vector(length = numberOfDays * 24)
for (hour in seq_len(numberOfDays * 24)) {
  days[hour] = hour / 24 
}

ggplot(dfDemand, aes(x = days, y = demands)) + geom_line(color = "dodgerblue") + 
  labs(title = "Texas Electricity Demands in 2018", x = "Days in 2018", y = "Net Demand of Texas (in MW)") 
```

Wind data

```{r}
dfDemand$Windless_Load = dfDemand$ERCOT.Load..MW - dfDemand$Total.Wind.Output..MW
windDemand = dfDemand$Windless_Load
windOutput = dfDemand$Total.Wind.Output..MW

ggplot(dfDemand, aes(x = days, y = windOutput)) + geom_line(color = "orange") + 
  labs(title = "Texas Wind Power Output in 2018", x = "Days in 2018", y = "MegaWatts")
```

Wind Power looks very sporadic while electricity demands seems to have a trend.


## Demand Prediction Strategy and Data Aggregation

For our independent variables we will use past week, past 2 days, past 1 day to predict the electiricty demand of tomorrow. i.e days to train on -7, -2, -1


```{r organize data}
daysToTrainOn = c(-7,-2,-1)
rangeOfDays = seq(-min(daysToTrainOn), numberOfDays, by = 1)

Y = NULL
for (day in rangeOfDays) {
  Y = rbind(Y, dfDemand$ERCOT.Load..MW[(day * 24): ((day + 1) * 24 - 1)])
}

X = NULL
for (day in rangeOfDays) {
  X_temp = cbind(t(dfDemand$ERCOT.Load..MW[(((day - 7)*24 +1)):((day - 7 + 1)*24)]),
            t(dfDemand$ERCOT.Load..MW[(((day - 2)*24) +1):((day - 2 + 1)*24)]),
            t(dfDemand$ERCOT.Load..MW[(((day - 1)*24) +1):((day - 1 + 1)*24)]))
  X = rbind(X, X_temp)
}
dim(X)
dim(Y)
```

After Organzing the data we will start making our train and test data.

## Prediction
